---
title: "IE 440-HW3"
author: "Alihan Zer - İrem Öztürk - Onur Can Aydın - Özlem Şenel"
output: html_document
---
In this homework, considering the following function:

f(x1, x2) = (5x1 − x2)^4 + (x1 − 2)^2 + x1 − 2x2 + 12

The minimum of this function are found using this methods:

1. Cyclic-Coordinate Search
2. Hook and Jeeves Method
3. Simplex Search

## Cyclic-Coordinate Search

In first set, epsilon=0.005 and x0=(0,30).
Here is the solution:

```{r cyclic1}
f <- function(x){
  (5*x[1] - x[2])^4 + (x[1]-2)^2 + x[1] - 2*x[2] + 12
}

e = matrix(nrow = 2, ncol = 2)
e[1,] = c(1,0)
e[2,] = c(0,1)

#selected alpha return
argmin_alpha<-function(x1,e,f){
  selected_alpha<-c()
  for (j in (1:2)) 
  {
    vec_all_alpha <- seq(-100, 100, by = 0.005)
    f_all_alpha <-c()
    for (i in 1:length(vec_all_alpha)) 
    {
      
      f_all_alpha[i] <-f(x1 + vec_all_alpha[i]*e[j,])
    }
    selected_alpha[j] <-vec_all_alpha[which.min(f_all_alpha)]
  }
  return(selected_alpha)
}


x1=c(0,30)


final_table <- data.frame(matrix(ncol = 8))
colnames(final_table) <- c("Iteration", "x(k)_1","x(k)_2", "f(x(k))" , "d(k)_x", "alpha(x)","x(k+1)_1", "x(k+1)_2")
cyclic <- function(x1,e,f,final_table){
  selected_alpha = c(0,0)
   k_fun=1
  while(1)
    {
     

    for ( j in (1:2))
      {
    
      selected_alpha<-argmin_alpha(x1,e,f)
       x2 <- x1+selected_alpha[j]%*%e[j,]
       
         
    if(k_fun==1){
     final_table[1,] = list(0, x1[1], x1[2], f(x1),NA, NA, NA, NA)
   }
   else{
     final_table[k_fun,] = list(k_fun-1, x1[1], x1[2], f(x1), (x2[2]-x1[2])+(x2[1]-x1[1]),(x2[2]-x1[2])+(x2[1]-x1[1]) , x2[1], x2[2])     
   }
       old_x1 <- x1
       x1<-x2
         k_fun=k_fun+1


    }
    if((abs(x1[2]-old_x1[2])+abs(x1[1]-old_x1[1]))<0.01)
       {
      print(final_table)
      print(paste("X coordinate of x(k): " ,x1[1], "Y coordinate of x(k): ", x1[2]))
      print(paste("f(x(k)): " ,f(x1)))
      break
      
       }

  }
}



cyclic(x1, e, f, final_table)


```

In second set, epsilon=0.005 and x0=(-10,10).
Here is the solution:

```{r cyclic2}
x1 = c(-10,10)
cyclic(x1, e, f, final_table)
```


## Hook and Jeeves Method

In first set, epsilon=0.005 and x0=(0,30).
Here is the solution:

```{r hook1}

f <- function(x){
  (5*x[1] - x[2])^4 + (x[1]-2)^2 + x[1] - 2*x[2] + 12
}
#selected alphas return
argmin_alpha<-function(x1,e,f){
  selected_alphas<-c()
  for (j in 1:2) 
  {
    vec_all_alpha <- seq(-100, 100, by = 0.005)
    f_all_alpha <-c()
    for (i in 1:length(vec_all_alpha)) 
    {
      f_all_alpha[i] <-f(x1 + vec_all_alpha[i]*e[j,])
    }
    selected_alphas[j] <-vec_all_alpha[which.min(f_all_alpha)]
  }
  return(selected_alphas)
}
argmin_alpha2<-function(x1,f,x3){
  alpha2_all <-seq(-100, 100, by = 0.05)
  f_pool2 <-vector()
  for (i in 1:length(alpha2_all)) 
  {
    f_pool2[i] <- f(x1 + alpha2_all[i]*x3)
  }
  alpha2 <- alpha2_all[which.min(f_pool2)]
  return(alpha2)
}

e = matrix(nrow = 2, ncol = 2)
e[1,] = c(1,0)
e[2,] = c(0,1)

x1 = c(0,30)

final_table <- data.frame(matrix(ncol = 11))
colnames(final_table) <- c("Iteration", "x(k)_1","x(k)_2", "f(x(k))" ,"x_temp_1","x_temp_2", "d(k)_1", "d(k)_2","alfa(k)", "x(k+1)_1","x(k+1)_2")
hook_jeeves <-function(x1, f,final_table){
 k_fun=1#function iteration
 
  while(1){
   
   selected_alphas<-argmin_alpha(x1,e,f)
    
   x2<-x1+selected_alphas%*%e
   #hold x_temp
   x3<-x2-x1
   
   alpha2<-argmin_alpha2(x1,f,x3)
   
   x_old<-x1
   x4 <- x1+alpha2*x3
   
   
   
   if(k_fun==1){
     final_table[1,] = list(0, x1[1], x1[2], f(x1), x2[1], x2[2], x3[1], x3[2],alpha2, NA, NA)
   }
   else{
     final_table[k_fun,] = list(k_fun-1, x1[1], x1[2],f(x1), x2[1], x2[2], x3[1], x3[2],alpha2, x4[1], x4[2])
     
   }
     
    x1 <- x4
     
   if(sqrt((x_old[1]-x1[1])^2+(x_old[2]-x1[2])^2) < 0.01){
     
     print(final_table)
      print(paste("X coordinate of x(k): " ,x1[1], "Y coordinate of x(k): ", x1[2]))
      print(paste("f(x(k)): " ,f(x1)))
     break
     
   }
   
    
    k_fun=k_fun+1
 }
} 


hook_jeeves(x1,f,final_table)


```

In second set, epsilon=0.005 and x0=(-10,10).
Here is the solution:

```{r hook2}

x1 = c(-10,10)

hook_jeeves(x1, f,final_table)


```


## Simplex Search

In first set, epsilon=0.005 and x1=(0.2,1), x2=(0,1) and x3=(0.4,0).
Here is the solution:

```{r simplex1}
f <- function(x){
  (5*x[1] - x[2])^4 + (x[1]-2)^2 + x[1] - 2*x[2] + 12
}

argmin <- function(f, x_list){
  helper <- c(f(x_list[[1]]), f(x_list[[2]]), f(x_list[[3]]))
  return(which.min(helper))
}

argmax <- function(f, x_list){
  helper <- c(f(x_list[[1]]), f(x_list[[2]]), f(x_list[[3]]))
  return(which.max(helper))
}

x1 <- c(0.2,1)
x2 <- c(0,1)
x3 <- c(0.4,0)
x_list <- list(x1,x2,x3)
final_table <- data.frame(matrix(ncol = 11))
colnames(final_table) <- c("Iteration", "x_bar: x", "x_bar: y" ,"x_h: x", "x_h: y", "x_l: x", "x_l: y", "x_new: x", "x_new: y", "f(x_new)", "Type")

simplex <- function(alpha, beta, gama, epsilon, f, x_list, final_table){
  iteration <- 1
  while(1){
    x_h <- x_list[[argmax(f, x_list)]]
    x_l <- x_list[[argmin(f, x_list)]]
    for(i in (1:3)){
      if(x_list[[i]]!=x_h && x_list[[i]]!=x_l){
        x_g <- x_list[[i]]
      }
    }
    x_c <- ((x_list[[1]] + x_list[[2]]+x_list[[3]])-x_h)/2
    
    
    x_r <- 2 * x_c - x_h
    x_new <- x_r
    type <- "R"
    
    if(iteration==1){
      final_table[1,] = list(0, x_c[1], x_c[2], x_h[1], x_h[2], x_l[1], x_l[2], NA, NA, NA, NA)
    }
    
    if(f(x_r)<f(x_l)){
      x_new <- (1+gama) * (x_c) - gama * x_h
      type <- "E"
    }
    
    else if(f(x_r)>=f(x_h)){
      x_new <- (1-beta) * x_c + beta * x_h
      type <- "C"
      }
    
    else if(f(x_g)<f(x_r) && f(x_r)<f(x_h)){
      x_new <- (1+beta) * x_c - beta * x_h
      type <- "C" 
    }
    
    x_list[[argmax(f, x_list)]] <- x_new
    
    helper_sum <- 0
    for(i in (1:3)){
      helper_sum <- helper_sum + (f(x_list[[i]]) -f(x_c))^2
    }
    
    final_table <- rbind(final_table, (list(iteration, x_c[1], x_c[2], x_h[1], x_h[2], x_l[1], x_l[2], x_new[1], x_new[2], f(x_new), type)))
    iteration <- iteration + 1
    
    if(sqrt(helper_sum/3)<epsilon){
      print(final_table)
      print(paste("X coordinate of x_bar: " ,x_c[1], "Y coordinate of x_bar: ", x_c[2]))
      print(paste("f(x_bar): " ,f(x_c)))
      
      break
    }
  }
} 

alpha = 1
beta = 0.5
gama = 2
epsilon = 0.005

simplex(alpha, beta, gama, epsilon, f, x_list, final_table)

```


In second set, epsilon=0.005 and x1=(1,10), x2=(20,0) and x3=(5,5).
Here is the solution:

```{r simplex2}


x1 <- c(1,10)
x2 <- c(20,0)
x3 <- c(5,5)
x_list <- list(x1,x2,x3)
simplex(alpha, beta, gama, epsilon, f, x_list, final_table)


```