---
title: "IE440 - HW1 - Group 6"
author: "Alihan Zer - Özlem Şenel - İrem Öztürk - Onur Can Aydın"
date: "06 11 2021"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

In this homework, we are asked to find a local minimum of this function between values a = -10 and b = 10:


f(x) = x^3 cos(x) sin(x) + 3x^2sin(x)-3x


Also, we are asked to find this value in several ways, which are:

- Bisection Method
- Golden Section Method
- Newton's Method
- Secant Method

```{r include=FALSE}
library(data.table)
f = function(x){
  (x^3) * cos(x*0.0174532925) * sin(x*0.0174532925) + 3 * (x^2) * sin(x*0.0174532925)-3*x
}

ff = expression((x^3) * cos(x*0.0174532925) * sin(x*0.0174532925) + 3 * (x^2) * sin(x*0.0174532925)-3*x)


```


```{r echo=FALSE, warning = FALSE}
library(ggplot2)
p <- ggplot(data = data.frame(x = 0), mapping = aes(x = x)) + stat_function(fun = f, col = "red") + xlim(-10,10) 
p + ggtitle("f(x) = x^3 cos(x) sin(x) + 3x^2sin(x)-3x")+ geom_point(x=2.893074, y= -6.191273)
```


This point shown in the graph is the global minimum point of this function. In all methods, x converges to this point in this graph.


```{r}

feq1 = function(x3,x2,x1, c){
  return(abs(x3-x2)/abs((x2-x1)^c))
}

feq2 = function(x3,x2,x1){
  return(log(abs(x2 - x1)) - log(abs(x3 - x2)))
}
```

Please note that, eq1 and eq2 symbolizes:



Equation_{1} = {|x^{(k+1)} - x^{(k)}|}{|x^{(k)} - x^{(k-1)}|}


Equation_{2} = -log|x^{(k+1)}-x^{(k)}| + log|x^{(k)} - x^{(k-1)}|


## Bisection Method

#### E = 0.005

```{r}

bisection = function(f, a, b, eps){
vec_k = c()
vec_a = c()
vec_b = c()
vec_x = c()
vec_fx = c()
vec_eq1 = c()
vec_eq2 = c()
counter = 0

while(b-a>=eps){
  x = (a+b)/2
  counter = counter + 1
  if(f(x+eps)<f(x)){
    a = x
  }
  else{
    b = x
  }
  vec_k[counter] = counter
  vec_x[counter] = x
  vec_a[counter] = a
  vec_b[counter] = b
  vec_fx[counter] = f(x)
}

for(i in (1:counter)){
  vec_eq1[i] = feq1(vec_x[i+1], vec_x[i], vec_x[i-1], c = 1)
  vec_eq2[i] = feq2(vec_x[i+1], vec_x[i], vec_x[i-1])
}
bisection_table = data.table(vec_k,vec_a, vec_b, vec_x, vec_fx, vec_eq1, vec_eq2)
 return(bisection_table) 
}

bt = bisection(f, -10, 10, 0.005)
colnames(bt) = c("Iteration", "a", "b", "x", "f(x)", "eq1", "eq2")
bt
```
This table shows all iterations. The epsilon value was chosen as 0.005.

Additionally, the optimal point and its value is:

```{r}
c(bt[.N, x], bt[.N, f(x)])
```
We can also choose a different epsilon value.

#### E = 0.001

```{r}
bt2 = bisection(f, -10, 10, 0.001)
colnames(bt2) = c("Iteration", "a", "b", "x", "f(x)", "eq1", "eq2")

bt2
```
The optimal point and its value is:

```{r}
c(bt2[.N, x], bt2[.N, f(x)])

```
#### E = 0.0005

```{r}
bt3 = bisection(f, -10 , 10, 0.0005)
colnames(bt3) = c("Iteration", "a", "b", "x", "f(x)", "eq1", "eq2")
bt3
```

The optimal point and its value is:

```{r}
c(bt3[.N, x], bt3[.N, f(x)])

```


## Golden Search

#### E = 0.005

```{r}
golden_section <-function(a,b,eps,f){
  
  x = b - 0.618 * (b-a)
  y = a + 0.618 * (b-a)
  fx = f(x)
  fy = f(y)
  counter=0
  vec_k<-c()
  vec_a<-c()
  vec_b<-c()
  vec_x<-c()
  vec_y<-c()
  vec_fx<-c()
  vec_fy <-c()
  vec_eq1 <- c()
  vec_eq2 <- c()
  
  while(b-a >= eps){
    
    counter = counter + 1
    
    if(fx > fy){
      a <- x
      x <- y
      fx <- fy
      y <- a + 0.618 * (b-a)
      fy <- f(y)
    }
    else{
      b <- y
      y <- x
      x <- b - 0.618 * (b-a)
      fy <- fx
      fx <- f(x)
    }
    
    vec_k[counter]=counter
    vec_a[counter]=a
    vec_b[counter]=b
    vec_x[counter]=x
    vec_y[counter]=y
    vec_fx[counter]=fx
    vec_fy[counter]=fy
  }
  vec_k <- append(vec_k, 0,after=0)
  vec_a <- append(vec_a, -10,after=0)
  vec_b <- append(vec_b, 10,after=0)
  vec_x <- append(vec_x, -2.36,after=0)
  vec_y <- append(vec_y, 2.36 ,after=0)
  vec_fx <- append(vec_fx, f(-2.36),after=0)
  vec_fy <- append(vec_fy, f(2.36),after=0)
  
  for(i in (1:counter)){
  vec_eq1[i] = feq1(vec_x[i+1], vec_x[i], vec_x[i-1], c = 1)
  vec_eq2[i] = feq2(vec_x[i+1], vec_x[i], vec_x[i-1])
  }
  
  vec_eq1 <- append(vec_eq1, NA ,after=0)
  vec_eq2 <- append(vec_eq2, NA ,after=0)
  
  
  dt_golden <- data.table(vec_k,vec_a, vec_b, vec_x,vec_y, vec_fx, vec_fy, vec_eq1, vec_eq2)
colnames(dt_golden) =  c ("Iteration", "a", "b", "x", "y", "f(x)", "f(y)", "eq1", "eq2")
  return(dt_golden)
  
}

gt = golden_section(-10, 10, 0.005, f)


gt
```
The optimal point and its value is:

```{r}
c(gt[.N, x], gt[.N, f(x)])

```
#### E = 0.001

```{r}
gt3 = golden_section(-10, 10, 0.001, f)
gt3
```
The optimal point and its value is:

```{r}
c(gt3[.N, x], gt3[.N, f(x)])

```
#### E = 0.0005

```{r}
gt2 = golden_section(-10, 10, 0.0005, f)
gt2


```

The optimal point and its value is:

```{r}
c(gt2[.N, x], gt2[.N, f(x)])

```

## Netwon's Method

#### E = 0.005

```{r}
derivative = function(k, ff){
  x = k
  eval(D(ff, 'x'))
}

second_derivative = function(k, ff){
  x = k
  eval(D(D(ff, 'x'), 'x'))
}

newton = function(x0, eps, f, ff){
  counter = 0
  vec_k = c()
  vec_xk = c()
  vec_fxk = c()
  vec_dfxk = c()
  vec_ddfxk = c()
  vec_eq1 = c()
  
  while(1){
    counter = counter + 1
    x1 = x0 - (derivative(x0, ff)/second_derivative(x0, ff))
    if(abs(derivative(x1, ff)) < eps){
      vec_k[counter] = counter
      vec_xk[counter] = x1
      vec_fxk[counter] = f(x1)
      vec_dfxk[counter] = derivative(x1, ff)
      vec_ddfxk[counter] = second_derivative(x1, ff)
      break
    }
    vec_k[counter] = counter
    vec_xk[counter] = x1
    vec_fxk[counter] = f(x1)
    vec_dfxk[counter] = derivative(x1, ff)
    vec_ddfxk[counter] = derivative(x1, ff)
    x0 = x1

    }
  for(i in (1:counter)){
    vec_eq1[i] = feq1(vec_xk[i+1], vec_xk[i], vec_xk[i-1], c = 2)
  }
  newton_table = data.table(vec_k, vec_xk, vec_fxk, vec_dfxk, vec_ddfxk,vec_eq1)
  colnames(newton_table) = c("Iteration", "x_k", "f(x_k)", "f'(x_k)", "f''(x_k)", "eq1")

 return(newton_table)
  }

nt = newton(10, 0.005, f, ff)
nt
```
The optimal point and its value is:

```{r}
c(nt[.N, x_k], nt[.N, f(x_k)])

```
#### E = 0.001

```{r}
nt2 = newton(10, 0.001, f, ff)
nt2
```

The optimal point and its value is:

```{r}
c(nt2[.N, x_k], nt2[.N, f(x_k)])

```


#### E = 0.0005

```{r}
nt3 = newton(10, 0.0005, f, ff)
nt3
```

The optimal point and its value is:

```{r}
c(nt3[.N, x_k], nt3[.N, f(x_k)])

```

## Secant Method

```{r include=FALSE}
derivative = function(k, ff){
  x = k
  eval(D(ff, 'x'))
}

x1_finder = function(x0, f){
  return(x0 - derivative(x0, ff)/derivative(x0, D(ff, 'x')))
}

```


#### E = 0.005

```{r}

secant = function(x0, x1, eps, f, ff){
   
    counter = 0
    vec_k = c()
    vec_xk = c()
    vec_fxk = c()
    vec_fdxk= c()
    vec_eq3 = c()
    while(1){
    counter = counter + 1
    x2 = x1 - ((derivative(x1, ff)) / ((derivative(x1, ff) - derivative(x0, ff)))) * (x1 - x0)
    
    if(abs(derivative(x2, ff)) < eps){
      vec_k[counter] = counter
      vec_xk[counter] = x2
      vec_fxk[counter] = f(x2)
      vec_fdxk[counter] = derivative(x2, ff)
      break
    }
    vec_k[counter] = counter
    vec_xk[counter] = x2
    vec_fxk[counter] = f(x2)
    vec_fdxk[counter] = derivative(x2, ff)
    x0 = x1
    x1 = x2
  
    }
  for(i in (1:counter)){
    vec_eq3[i] = (abs(vec_xk[i+1] - vec_xk[i])) / abs(vec_xk[i] - vec_xk[i-1])^1.618
  }
  secant_table = data.table(vec_k, vec_xk, vec_fxk, vec_fdxk, vec_eq3)
  colnames(secant_table) = c("Iteration", "x_k", "f(x_k)", "f'(x_k)",  "eq1")

 return(secant_table)
}




st = secant(10, x1_finder(10, f), 0.005, f, ff)
colnames(st) = c("Iteration", "x_k", "f(x_k)", "f'(x_k)",  "eq1")
st
```
The optimal point and its value is:

```{r}
c(st[.N, x_k], st[.N, f(x_k)])

```


#### E = 0.001

```{r}
st2 = secant(10,x1_finder(10, f), 0.001, f, ff)
st2
```

The optimal point and its value is:

```{r}
c(st2[.N, x_k], st2[.N, f(x_k)])

```

#### E = 0.0005

```{r}
st3 = secant(10,x1_finder(10, f), 0.0005, f, ff)
st3
```

The optimal point and its value is:

```{r}
c(st3[.N, x_k], st3[.N, f(x_k)])

```
