---
title: "IE440-HW4"
author: "Alihan Zer - İrem Öztürk - Onur Can Aydın - Özlem Şenel"
output: html_document
---


In this homework, considering the following function:

f(x1, x2) = (5x1 − x2)^4 + (x1 − 2)^2 + x1 − 2x2 + 12


The minimum of this function are found using this methods:

1. Newton’s Method (with exact line search)
2. DFP (Davidon - Fletcher - Powell) Method
3. BFGS (Broyden - Fletcher - Goldfarb - Shanno) Method



## Newton’s Method (with exact line search)

For the function, epsilon=0.005 and x0=(1,10).
Here is the solution:


```{r newton1}
f1 <- function(x){
  (5*x[1] - x[2])^4 + (x[1]-2)^2 + x[1] - 2*x[2] + 12
}
f1_by_x1<-function(x){
  4 * (5 * (5 * x[1] - x[2])^3) + 2 * (x[1] - 2) + 1
}

f1_by_x1x1<-function(x){
  300 * (5 * x[1] - x[2])^2 + 2
}

f1_by_x2<-function(x){
  -(4 * (5 * x[1] - x[2])^3 + 2)
}

f1_by_x2x2<-function(x){
  12 * (5 * x[1] - x[2])^3
}

f1_by_x1x2<-function(x){
  -60 * (5 * x[1] - x[2])^2
}

f2 <- function(x){
  100 * (x[2] - x[1]^2)^2 + (1-x[1])^2
}
f2_by_x1<-function(x){
  -(2 * (1 - x[1]) + 100 * (2 * (2 * x[1] * (x[2] - x[1]^2))))
}

f2_by_x1x1<-function(x){
  1600 * (x[1]^2) * (x[2] - x[1]^2) + 400 * (x[2] - x[1]^2)^2 + 2
}
f2_by_x2<-function(x){
  100 * (2 * (x[2] - x[1]^2))
}

f2_by_x2x2<-function(x){
  400 * (x[2] - x[1]^2)
}

f2_by_x1x2<-function(x){
  (-8) * x[1] * (x[2] - x[1]^2)
}

grad_f1<-function(f1,x){
  grad_f = c(1,1)
  grad_f[1]=f1_by_x1(x)
  grad_f[2]=f1_by_x2(x)
  return(grad_f)
  
}

grad_f2<-function(f2,x){
  grad_f = c(1,1)
  grad_f[1]=f2_by_x1(x)
  grad_f[2]=f2_by_x2(x)
  return(grad_f)
  
}

second_grad_f1<-function(f1,x){
  grad_f = matrix(1:4, nrow = 2, ncol = 2)
  grad_f[1,1]=f1_by_x1x1(x)
  grad_f[1,2]=f1_by_x1x2(x)
  grad_f[2,1]=f1_by_x1x2(x)
  grad_f[2,2]=f1_by_x2x2(x)
  return(grad_f)
}

second_grad_f2<-function(f2,x){
  grad_f = matrix(1:4, nrow = 2, ncol = 2)
  grad_f[1,1]=f2_by_x1x1(x)
  grad_f[1,2]=f2_by_x1x2(x)
  grad_f[2,1]=f2_by_x1x2(x)
  grad_f[2,2]=f2_by_x2x2(x)
  return(grad_f)
}



argmin_alpha<-function(x1,d,f){
  selected_alphas<-c(1)
  
  vec_all_alpha <- seq(-100, 100, by = 0.005)
  f_all_alpha <-c()
  for (i in 1:length(vec_all_alpha)) 
  {
    f_all_alpha[i] <-f(x1 + vec_all_alpha[i]*d)
  }
  selected_alphas <-vec_all_alpha[which.min(f_all_alpha)]
  
  return(selected_alphas)
}


final_table <- data.frame(matrix(ncol = 9))
colnames(final_table) <- c("Iteration", "x(k)_1","x(k)_2", "f(x(k))" ,"d_1","d_2", "alpha", "x(k+1)_1","x(k+1)_2")




Newton1 <- function(f1,x1,final_table,epsilon){
  k_fun=1
  d <- list()
  alpha <- c()
  x <- list()
  x[[1]] = x1
  while(1){
    d[[k_fun]]<- (-1)*solve(second_grad_f1(f1,x[[k_fun]])) %*% grad_f1(f1,x[[k_fun]])
    alpha[k_fun]<-argmin_alpha(x[[k_fun]],d[[k_fun]],f1)
    
    x[[k_fun+1]] <- x[[k_fun]] + alpha[k_fun] * d[[k_fun]]
    
    
    
    final_table[k_fun,] = list(k_fun-1, x[[k_fun]][1], x[[k_fun]][2],f1(x[[k_fun]]), d[[k_fun]][1], d[[k_fun]][2], alpha[k_fun], x[[k_fun+1]][1], x[[k_fun+1]][2])
    
    k_fun = k_fun + 1
    
    a<-grad_f1(f1,as.vector(t(x[[k_fun]])))
    if(sqrt(a[1]^2+a[2]^2)<epsilon){
      print(final_table)
      print(paste("X coordinate of x_bar: " ,x[[k_fun]][1], "Y coordinate of x_bar: ", x[[k_fun]][2]))
      print(paste("f(x_bar): " ,f1(x[[k_fun]])))
      break
      
    }
    
    else if(k_fun==50){
      print(final_table)
      print(paste("X coordinate of x_bar: " ,x[[k_fun]][1], "Y coordinate of x_bar: ", x[[k_fun]][2]))
      print(paste("f(x_bar): " ,f1(x[[k_fun]])))
      break
    }

  }
  
}

Newton2 <- function(f2,x1,final_table,epsilon){
  k_fun=1
  d <- list()
  alpha <- c()
  x <- list()
  x[[1]] = x1
  while(1){
    d[[k_fun]]<- (-1)*solve(second_grad_f2(f2,x[[k_fun]])) %*% grad_f2(f2,x[[k_fun]])
    alpha[k_fun]<-argmin_alpha(x[[k_fun]],d[[k_fun]],f2)
    
    x[[k_fun+1]] <- x[[k_fun]] + alpha[k_fun] * d[[k_fun]]
    
    
    
    final_table[k_fun,] = list(k_fun-1, x[[k_fun]][1], x[[k_fun]][2],f2(x[[k_fun]]), d[[k_fun]][1], d[[k_fun]][2], alpha[k_fun], x[[k_fun+1]][1], x[[k_fun+1]][2])
    
    k_fun = k_fun + 1
    
    a<-grad_f2(f2,as.vector(t(x[[k_fun]])))
    if(sqrt(a[1]^2+a[2]^2)<epsilon || alpha[[k_fun-1]] == 0){
      print(final_table)
      print(paste("X coordinate of x_bar: " ,x[[k_fun]][1], "Y coordinate of x_bar: ", x[[k_fun]][2]))
      print(paste("f(x_bar): " ,f1(x[[k_fun]])))
      break
      
    }
    
    else if(k_fun==50){
      print(final_table)
      print(paste("X coordinate of x_bar: " ,x[[k_fun]][1], "Y coordinate of x_bar: ", x[[k_fun]][2]))
      print(paste("f(x_bar): " ,f1(x[[k_fun]])))
      
      break
    }
    
  }
  
}


x1<-c(1,10)
epsilon=0.005

Newton1(f1,x1,final_table,epsilon)
```


## DFP (Davidon - Fletcher - Powell) Method

For the function, epsilon=0.005 and x0=(1,10).
Here is the solution:

```{r dfp1}
f_1 <- function(x){
  (5*x[1] - x[2])^4 + (x[1]-2)^2 + x[1] - 2*x[2] + 12
}
f_1_by_x1<-function(x){
  4 * (5 * (5 * x[1] - x[2])^3) + 2 * (x[1] - 2) + 1
}
f_1_by_x2<-function(x){
 -(4 * (5 * x[1] - x[2])^3 + 2)
}
f_1_exp<-expression((5*x1 - x2)^4 + (x1-2)^2 + x1 - 2*x2 + 12)


f_2 <- function(x){
100 * (x[2] - x[1]^2)^2 + (1-x[1])^2
}
f_2_by_x1<-function(x){
  -(2 * (1 - x[1]) + 100 * (2 * (2 * x[1] * (x[2] - x[1]^2))))
}
f_2_by_x2<-function(x){
  100 * (2 * (x[2] - x[1]^2))
}
f_2_exp<-expression(100*(x2-x1^2)^2 + (1-x1)^2)

grad_f1<-function(f1,x){
  grad_f = c(1,1)
  grad_f[1]=f1_by_x1(x)
  grad_f[2]=f1_by_x2(x)
  return(grad_f)

}

grad_f2<-function(f2,x){
  grad_f = c(1,1)
  grad_f[1]=f2_by_x1(x)
  grad_f[2]=f2_by_x2(x)
  return(grad_f)

}


library(data.table)


A_calc <- function(p,q){
  numerator = p %*% t(p)
  denominator = t(p) %*% p
  return(numerator / as.numeric(denominator))
}

B_calc <- function(q, H){
  numerator = H %*% q %*% t(q) %*% H
  denominator = t(q) %*% H %*% q
  return(-(numerator/as.numeric(denominator)))
}


euc_dist = function(d){
  return(sqrt(d[1]^2 + d[2]^2))  
}


argmin_alpha<- function(x,d,f){
  selected_alpha<-c()
  vec_all_alpha <- seq(0.005, 200, by = 0.005)
  f_all_alpha <-c()
   for (i in 1:length(vec_all_alpha)){
     
    f_all_alpha[i] <-f(x + vec_all_alpha[i]*d)
     }
  selected_alpha <-vec_all_alpha[match(min(f_all_alpha), f_all_alpha)]
  
  return(selected_alpha)
}

DFP <- function(f, epsilon){
  if(identical(deparse(f), deparse(f_1))){
    gradd = grad_f1
  }
  else{
    gradd = grad_f2
  }
  
final_table <- data.frame(matrix(ncol = 9))
colnames(final_table) <- c("Iteration(k)", "x1(k)","x2(k)", "fx(k)" ,"d(k):X1", "d(k):x2", "alpha(k)", "x1(k+1)","x2(k+1)")
x = list()
H = list()
x[[1]] = c(0,0)
k = 1
H[[k]] = diag(2)

d = list()
alpha = c()
p = list()
q = list()
A = list()
B = list()

while(1){
  d[[k]] =  -H[[k]] %*% gradd(f, x[[k]]) 
  if(euc_dist(d[[k]]) < epsilon){
    break
  }
 
  else{
    alpha[k] = argmin_alpha(x[[k]], as.vector(d[[k]]), f)
    x[[k + 1]] = x[[k]] + alpha[k] * d[[k]]
    p[[k]] = x[[k+1]] - x[[k]]  
    q[[k]] = gradd(f, x[[k+1]]) - gradd(f, x[[k]])
    A[[k]] = A_calc(p[[k]], q[[k]])
    B[[k]] = B_calc(q[[k]], H[[k]])
    H[[k+1]] = H[[k]] + A[[k]] + B[[k]]
    
    final_table = rbind(final_table, list(k-1, x[[k]][1], x[[k]][2] ,f(x[[k]]), d[[k]][1],d[[k]][2], alpha[k], x[[k+1]][1],x[[k+1]][2] ))
    k = k + 1
    
  }
  
}
print(paste("x*=", x[[k-1]]))
print(paste("f(x*)=", f(x[[k-1]])))
rownames(final_table) = 0:(nrow(final_table)-1)
return(final_table[-1,])
}
table_dfp_1= DFP(f_1,0.005)
table_dfp_1

```


## BFGS (Broyden - Fletcher - Goldfarb - Shanno) Method

For the function, epsilon=0.005 and x0=(0,0).
Here is the solution:

```{r bfgs1}

A_calc_2 <- function(p,q,H){
  
  numerator1 =  as.numeric(1 + t(q) %*% H %*% q)
  denominator1 = as.numeric(t(q) %*% p)
  
  numerator2 = p %*% t(p)
  denominator2= as.numeric(t(p) %*%q)
  
  return((numerator1 / denominator1)*(numerator2/denominator2))
}

B_calc_2 <- function(p,q, H){
  
  numerator = (p %*% t(q) %*% H )+ (H %*% q %*% t(p))
  denominator = as.numeric(t(q) %*% p)
  return(-(numerator/denominator))
}




BFGS <- function(f, epsilon){
    if(identical(deparse(f), deparse(f_1))){
    gradd = grad_f1
  }
  else{
    gradd = grad_f2
  }

  final_table <- data.frame(matrix(ncol = 9))
  colnames(final_table) <- c("Iteration(k)", "x1(k)","x2(k)", "fx(k)" ,"d(k):X1", "d(k):x2", "alpha(k)", "x1(k+1)","x2(k+1)")
  x = list()
  H = list()
  x[[1]] = c(0,0)
  k = 1
  H[[k]] = diag(2)
  
  d = list()
  alpha = c()
  p = list()
  q = list()
  A = list()
  B = list()
  
  while(1){
    d[[k]] =  -H[[k]] %*% gradd(f, x[[k]]) 
    if(euc_dist(d[[k]]) < epsilon){
      break
    }
    
    else{
      alpha[k] = argmin_alpha(x[[k]], as.vector(d[[k]]), f)
      x[[k + 1]] = x[[k]] + alpha[k] * d[[k]]
      p[[k]] = x[[k+1]] - x[[k]]  
      q[[k]] = gradd(f, x[[k+1]]) - gradd(f, x[[k]])
      A[[k]] = A_calc_2(p[[k]], q[[k]], H[[k]])
      B[[k]] = B_calc_2(p[[k]], q[[k]], H[[k]])
      H[[k+1]] = H[[k]] + A[[k]] + B[[k]]
      
      final_table = rbind(final_table, list(k-1, x[[k]][1], x[[k]][2] ,f(x[[k]]), d[[k]][1],d[[k]][2], alpha[k], x[[k+1]][1],x[[k+1]][2] ))
      k = k + 1
      
    }
    
  }

  print(paste("x*=", x[[k-1]]))
  print(paste("f(x*)=", f(x[[k-1]])))
  rownames(final_table) = 0:(nrow(final_table)-1)
  return(final_table[-1,])
}



table_bfgs_1= BFGS(f_1,0.005)
table_bfgs_1

```

